% -*- root: ../main.tex -*-
%!TEX root = ../main.tex
% this file is called up by main.tex
% content in this file will be fed into the main document
% vim:textwidth=80 fo=cqt

In this  section, the performance  of the  basic \gls{spm} is  discussed through
simulation  and  by comparison  against  a  standard \gls{dfn}  benchmark  model
incorporating the full \gls{p2d} dynamics.

\subsection{Cell Parameters}
\input{4/cell_params_table}

\Cref{tbl:lcoSimParamsSPMp2d} lists  the simulation  parameters of  an \gls{lco}
cell  whose positive  and negative  electrodes are  \ch{LiCoO_2} and  \ch{LiC_6}
respectively.  The  electrolyte in  this  system  consists of  \ch{LiPF_6}  salt
in  a  solution of  \gls{ec}/\gls{dmc}/\gls{emc}  in  a  1:1:1 ratio.  The  vast
majority  of electrochemical  parameters, \viz{}  the geometric,  thermodynamic,
kinetic  and   transport  properties  of   the  cell  have  been   sourced  from
Subramanian~\etal{}~\cite{Subramanian2009}.

The simulation parameters that are exclusively applicable to the \gls{p2d} model
are shown with a shaded highlighting in~\cref{tbl:lcoSimParamsSPMp2d}. It can be
seen  that only  a subset  of the  isothermal \gls{dfn}  model's parameters  are
required for  the the  \gls{spm}. In particular,  the~\gls{spm} has  seven fewer
parameters  per  electrode. Furthermore,  all  six  physical parameters  of  the
separator  material required  in the  \gls{dfn} model  are ignored  in \gls{spm}
computations. Neglecting Arrhenius type temperature dependence of parameters and
their  corresponding activation  energies, the  basic \gls{spm}  facilitates the
ability to afford physics-based  modelling capabilities with \emph{twenty} fewer
parameters  than  the equivalent  isothermal  \gls{p2d}  model. With  the  naive
assumption of equal  parametrisation effort per physical  property, this implies
an  exact \SI{50}{\percent}  reduction in  parametrisation requirements  for the
basic \gls{spm} compared its \gls{dfn} counterpart.


The volume  fraction of the electrode and separator
materials are computed as $1 - \varepsilon_j - \varepsilon_{\text{fi}_j}$, where
$\varepsilon_j$  is the  electrolyte-phase volume  fraction within  each of  the
three regions of the electrode.

% \rowcolor{imperiallightgray} Material vol.\ fraction, $\varepsilon_\sj$                         & \tnote{d}\num{0.590}    & \tnote{d}\num{0.276}     & \tnote{d}\num{0.482}                                 \\

In the interest of reproducibility of  results, the design decisions made by the
author on  a few important aspects  of setting up the  simulations are discussed
next.

\subsection{Simulation Setup}

Smaller parametrisation requirements  is only half the story  towards speed The.
cells lower cutoff voltage is chosen to be \SI{2.5}{V}                         .
