% -*- root: ../../main.tex -*-
%!TEX root = ../../main.tex
% vim:textwidth=80 fo=cqt

% As discussed  in \cref{ch:intro},  there is  a growing  need to  replace replace
% fossil  fuel-based  vehicles  with   cleaner  alternatives  wherein  Lithium-ion
% batteries hold  much promise.  The next generation  of battery  powered vehicles
% shall  require  higher  energy  density  batteries  without  compromising  power
% density. Furthermore,  the costs of the  on-board battery pack shall  have to be
% considerably lowered.

% Researchers in the field have aimed to bring about improvements to the incumbent
% lithium ion battery technology by focussing on various aspects.

% Modifications are  attempted  through various  approaches i)  through
% fundamental material advances [  2 , 3 ], ii) new chemistries [  4 ], iii) novel
% cell designs and  manufacturing techniques [ 5 ], iv)  system design or reducing
% the costs of assembly [ 6 , 7  ], and v) improved controller design for advanced
% battery management systems [8].

\subsection{Introduction}

In  sharp  contrast to  the  cornucopia  of  published literature  dealing  with
reduced order modelling of cells (see \cref{sec:classificationscheme}), there is
currently  a relative  paucity  of  prior art  that  discusses model-based  cell
design. Albeit the relevant pool of  knowledge is presently sparse, this section
nevertheless aims to critically evaluate it.

At the outset, it is important to  clarify this thesis author's general views on
model-based design. It is  also important to state why this  topic was chosen as
an  aspect of  this  thesis. In  the recent  decades,  industries pertaining  to
different  walks of  life,  cutting across  any one  discipline,  have begun  to
establish a model-led product/process development culture. For instance, a broad
overview of  industrial research policy by  Thomke~\cite{Thomke1998} towards the
turn of  the last  millennium provides quantifiable  evidence that  a simulation
based design  approach in the automotive  industry has had a  positive impact in
the  crash-worthiness of  resulting  vehicular  designs. At  a  high level,  the
salient benefits of any model-based design include ---
\begin{enumerate*}[label=\itshape\alph*\upshape)]
    \item reducing the number of design iterations thereby speeding up the time to a production-ready prototype, and
    \item facilitating improved understanding of the variables influencing design that is gained by formalising empirical/ad-hoc knowledge through modelling.
\end{enumerate*}
Furthermore, it  is a well-known fact  that computer-time is cheaper  than human
time. Therefore, with a simulation-oriented design approach, it is often cheaper
to  explore  Monte~Carlo-like  design  scenarios using  a  computer  model  than
iterating over a series of rudimentary  prototypes in the lab. Finally, once the
results predicted  by the  model have  stabilised enough  to satisfy  the design
specifications, prototypes closely  matching the final design  objectives can be
realised, thereby reducing the overall lead time to the market.

The   aforementioned    views   of    the   thesis    author   is    echoed   by
Becker~\etal~\cite{Becker2005}  who present  a  persuasive view  that, since  it
shall be  mandatory to have  a deeper understanding  of the simulation  tools in
order  to successfully  employ a  model-led  design, this  can trigger  sweeping
changes percolating  into the  very core  of the  problem-solving culture  in an
organisation. Although relatively  at its infancy when it comes  to cell design,
model-based designs have been applied at  the battery pack-level in the past and
is therefore not a newcomer to the battery industry in general. In the middle of
the last  decade, a clarion  call to the industry  to adopt simulation  tools in
battery  engineering was  issued  by Spotnitz~\cite{Spotnitz2005}.  In the  said
article,  the author  questioned  the anachronistic  industry  trend of  relying
heavily on `making and testing' rather than aiming to understand the fundamental
governing equations  and principles  of a  battery and  using this  know-how for
design. Spotnitz~\cite{Spotnitz2005}  further argues  that using  \glspl{pbm} of
batteries could provide  reliable understanding of their behaviour,  and that as
the  understanding of  the  community steadily  grows, it  could  bring about  a
significant speed-up of battery development.

The nature and  scope of `model-based design' as intended  by this thesis author
needs to  be clarified. Since  this thesis focuses exclusively  on physics-based
modelling,  the  term  `model'  as  used  here  pertains  to  \glsfmtlongpl{pbm}
(\glsfmtshortpl{pbm}) and  not \glspl{ecm} or  any other type of  battery models
including other empirical/ad-hoc models such as surrogate models. Therefore, the
survey of  literature here  does not  include any  prior efforts  on model-based
design that lie outside of this scope.

Among  the published  set of  literature, there  is evidence  that computational
modelling  has  been   successfully  used  to  facilitate   the  development  of
novel   energy   storage   materials.   The  review   article   by   Islam   and
Fisher~\cite{Islam2014} provides an  overview of the use  of computer simulation
into  gaining  a  deeper insight  into  the  working  of  new types  of  cathode
materials. Meng  and Arroyo-de~Domp~\cite{Meng2009}  also surveyed the  topic of
using  computational tools  for the  design and  optimisation of  energy storage
materials. The  computer models  referred to  in these  review articles  are not
volume-averaged  models  operating at  the  cell-level,  but detailed  ab-initio
models  constructed using  techniques such  as Density  Functional Theory~(DFT).
While  it is  heartening  to  see such  comprehensive  studies of  computational
techniques being applied  to energy storage, there are two  distinct reasons why
the body of  research reviewed in the aforementioned articles  do not align with
the goals of  this thesis. Firstly, in the works  reviewed in the aforementioned
articles, computer simulation is used  primarily to enhance researchers' current
understanding of these materials that can help to develop the next generation of
energy  storage  components. The  computational  methods  presented therein  are
\emph{not}  directly employed  as  design tools.  Secondly,  these articles  use
computational  modelling to  study structural  properties at  the meso  and nano
scales.  While  these  are  of  utmost importance  to  researchers  involved  in
synthesising prototypes of next generation of energy storage materials, they are
less relevant for production-ready cell-designs  at scale. Since this thesis has
a  strong  focus on  providing  readily  applicable  solutions to  industry  for
incumbent lithium  ion chemistries,  it was  decided not  to pursue  the methods
discussed in the aforementioned works for the design studies discussed herein.

A      holistic      computational      screening     was      performed      by
Sendek~\etal~\cite{Sendek2017}  to  study  the  suitability  of  12831~candidate
materials for their  suitability as solid state  electrolytes in electrochemical
cells. These  authors cite the  same rationale  of this thesis  author \ie~rapid
prototyping, as the  motivation behind this model-based  design simulations. The
study  helped to  narrow the  initial candidate  pool down  to 21~viable  family
structures. This effort  serves as a concrete example in  this literature survey
wherein computer simulation is directly used  as a design tool for components of
a lithium  ion cell. The  scope of  the aforesaid work  falls into the  realm of
material  synthesis  and applied  physics  ---  topics outside  the  educational
background and  expertise of  this thesis author.  Nevertheless, the  success of
this effort  strongly motivates  the case for  performing a  computational study
of  similar scale,  wherein  instead  of screening  out  thousands of  candidate
materials, the framework presented in \cref{ch:modelbaseddesign} computationally
screens out thousands of layer configurations within a pouch cell.

Curiously,  the  use  of  volume-averaged   models  for  design  simulations  of
\glspl{pbm} at  the cell-scale has not  yet gained sufficient traction.  This is
despite the prevalence of the  popular \gls{p2d} implementation of the \gls{dfn}
model that  in applications such  as degradation analysis  and state-estimation.
Ramadesigan~\etal~\cite{Ramadesigan2012} postulate  that this slow  uptake could
be attributed to the computational challenges presented by the complex reaction,
diffusion and kinetics of lithium ion  cells occurring over different length and
time scales. A few design efforts using electrochemical models at the cell scale
have nevertheless been reported in literature, which are examined next.

\subsection{State of the art in cell-scale model-led design optimisations}

The pioneering  work by Newman~\cite{Newman1995}  was the  first of its  kind to
develop a scheme to optimise cell  design based on an electrochemical model. The
two parameters optimised  were electrode thicknesses and  porosities. This study
makes the assumption that electrode kinetics are fast relative to diffusion, and
furthermore ignores local concentration  gradients. A reaction-zone model, which
considers that  reactions occur  in a  narrowly confined area,  was used  as the
underlying  model. The  separator  thickness was  assumed to  be  fixed and  the
specific energy of the cell was  maximised. However, the aforesaid work does not
impose any constraints  on the extent of  specific power that can  be drawn from
the cell. Nevertheless,  this seminal effort provided the key  guidance to other
researchers that electrode  thicknesses and porosities can be  considered as the
critical design  variables to be optimised  for in a cell  design. Building upon
the foundation  laid by the aforementioned  study, a few other  model-led design
efforts have  been published. However,  in order  to provide a  focussed review,
only a  subset of  the prior art  that is deemed  pertinent for  the model-based
design of  \emph{pouch} cells  for \emph{automotive} applications  is considered
here.

Arora~\etal~\cite{Arora1999}  used an  electrochemical  model  to inform  design
decisions on cell parameters such as particle size, electrode thickness and mass
ratio. Furthermore, the model thus obtained was used to optimise the cell design
against the  risk of Lithium  plating. Cells  are susceptible to  plating during
fast charging and hence a key aspect to consider in the cell design for electric
vehicles. This  thesis author considers  this contribution  to be a  vanguard in
model-based  design  for cells  that  is  potentially applicable  for  vehicular
applications. Nevertheless, the aspect of stacking layers inside pouch cells and
how this influences the optimisation  of the aforementioned design variables has
not been studied.

Xue~\etal~\cite{Xue2013}  presented a  simulation based  design study  wherein a
cell design  was optimised using  a gradient-based algorithm.  Specifically, the
said  work dealt  with  a  numerical framework  for  providing automated  design
outputs for maximising  the cell's energy density whilst  meeting specific power
density  requirements. This  criterion  appears  to be  highly  relevant in  the
context of cell design for electric vehicles and hence, was adopted as the basis
for the layer optimisation framework presented in \cref{ch:modelbaseddesign}. In
the aforementioned  work, various power  levels were also tried.  However, these
power requirements were  computed indirectly rather than  a direct reformulation
of the underlying \gls{p2d} model to accept power density inputs.

% reformulated to account for applied power density. Particularly notable is the
% absence of relevant boundary conditions to the governing \glspl{pde}.


The  design   study  by  Xue~\etal~\cite{Xue2014}  represents   a  rare  example
of  a  model-led  design  attempt  performed  in  the  backdrop  of  electrified
transportation.  In  the  said  work, an  electrochemical  cell  model  \viz~the
\gls{p2d} model was successfully adapted to perform the design optimisation of a
\gls{phev}. Although the underlying model operates at the cell-scale, this study
considers  a pack-level  optimisation through  a series-parallel  combination of
cells. Furthermore,  an advanced  numerical algorithm  in the  form of  a hybrid
solver  was  used which  used  a  unique  scheme  of employing  a  gradient-free
optimiser  in conjunction  with a  gradient-based optimiser.  Albeit a  standout
example  in the  context  of  vehicular application,  the  aforesaid work  falls
slightly outside the  scope of this thesis wherein the  design study is strictly
confined to  be at the  cell level. The  aforementioned work takes  into account
pack-level constraints such as safety limits as well as energy and power levels.
Furthermore,  these authors  even accounted  for important  details such  as the
presence of layers within each cell in their design optimisation process.

The  system-level   constraints  considered  in   Xue~\etal~\cite{Xue2014}  were
instrumental  in  informing  the  choice  of constraints  in  the  design  study
presented in \cref{ch:modelbaseddesign}. However, in the aforementioned article,
the translation from pack-level into  cell-level quantities does not account for
secondary effects \ie~the  influence of layers on pack mass  (a quantity that is
optimised in the said publication) is not studied. Simulations conducted as part
of the  work reported in  \cref{ch:modelbaseddesign} revealed that  such effects
are non-negligible.

% Furthermore, although  not
% explicitly mentioned,  it is  assumed that, despite  layer counts  being changed
% during  optimisation,  the  electrode  thicknesses are  held  constant,  thereby
% violating  key principles  such as  active  material capacity  balancing of  the
% electrodes.

A critical issue in the approach  by Xue~\etal~\cite{Xue2014} is that the number
of design variables were excessively large which necessitated the use of complex
numerical  algorithms.  The  said  work  assumes  that  the  cell  design  shall
be  an  integral  part  of  the  pack  optimisation.  However,  in  this  thesis
author's  understanding,  it  appears  that in  real-world  designs,  cells  are
sourced from a  specialised manufacturer. Furthermore, these  cells are designed
independently to merely  adhere to certain specifications~\cite{Maksimovic2012}.
Pack configuration  decisions such  as the choice  of number  of series-parallel
modules in  the vehicular pack are  undertaken at a system  level in conjunction
with the  rest of the  drivetrain specifications by vehicle  manufacturers. This
natural separation of the real-world paradigm  helps to decouple the cell design
from  the pack  design thereby  drastically reducing  the number  of degrees  of
freedom and simplifying  the optimisation task. This is a  key assumption in the
design optimisation study presented in \cref{ch:modelbaseddesign}.

A  comprehensive multi-objective  optimisation for  optimal design  of batteries
was  recently  proposed  by  Changhong  Liu  and~Lin  Liu~\cite{Liu2017b}.  This
optimisation  involved a  multiphysics model  of the  cell wherein  a number  of
design variables  such as electrode  thicknesses, porosities and  particle sizes
were considered. The optimisation objectives were to maximise the specific power
and specific energy  during discharge as well as to  minimise capacity loss. The
model  was solved  by  using a  genetic algorithm.  While  this contribution  is
certainly  the first  of  its kind  to use  a  formal mathematical  optimisation
framework for cell design, in the  opinion of this thesis author, the complexity
of the problem is excessively  high --- both mathematically and computationally.
For  instance, with  three design  variables considered,  it requires  a careful
interpretation  of the  resulting  pareto front  to  distinguish their  relative
importance. While  prima~facie this is not  an issue, the question  of resolving
the pareto front for cells with  different parameter sets needs to be addressed.
The optimisation algorithm  used is an exotic  genetic algorithm to be  run on a
parallel cluster. From a near-term industrial application viewpoint, this scheme
is less attractive  since real-world constraints, such  as supply-chain dictated
raw-material  sourcing, typically  restrict  the number  of  degrees of  freedom
available  for  optimisation.  Nevertheless,  this scheme  is  valuable  from  a
long-term research perspective and merits attention by researchers interested in
model-based design.

From the  body of the  relevant literature surveyed here,  it is clear  that the
issue of optimally  stacking up layers within  a pouch cell, has  not been dealt
with in a  systematic manner. Pouch cells  are the most common type  of cells in
automotive  applications.  Therefore,  optimising  their design  shall  yield  a
beneficial improvement to the overall driving range and fast charging capability
of electric  vehicles while simultaneously  helping to increase the  lifetime of
their  battery  packs.  With  these goals  in  mind,  \cref{ch:modelbaseddesign}
presents  a mathematical  framework for  a model-based  optimal design  of pouch
cells.


