% -*- root: ../../main.tex -*
%!TEX root = ../../main.tex
% vim:textwidth=80 fo=cqt conceallevel=0

\newpage
\thispagestyle{plain} % empty
\mbox{}
\graphicspath{{chapters/layer_opt/figures/}}
% ----------------------- contents from here ------------------------

\chapter[Model-based Design of Pouch Cells]{%
    Model-based  Design Of  Pouch Cells\footnote{\textbf{Attribution  of content}  \
        The  groundwork  for converting  the  existing  computer code  (LIONSIMBA~v1.0x)
        into  a suitable  form  for  layer optimisation  was  initiated  by this  thesis
        author, \mbox{Krishnakumar  Gopalakrishnan}. However, with the  exception of the
        spectral scheme,  for which \mbox{Krishnakumar Gopalakrishnan}  was responsible,
        and  the  zero-dimensional  thermal  model for  which  \mbox{Ian  D.\  Campbell}
        (PhD  student,  Imperial  College  London)  was  responsible,  the  advancements
        inherent to the  enhanced computer software (LIONSIMBA~v2.0) were  made in equal
        parts by  \mbox{Krishnakumar Gopalakrishnan} and \mbox{Ian  D.\ Campbell}. These
        advancements would not have been  possible without the contributions and support
        of \mbox{Dr~Davide M.~Raimondo} (Associate  Professor, University of Pavia) who
        served as  an unofficial supervisor for  the work reported in  this chapter. The
        concept  of layer  reconfiguration for  energy  and power  trade-off, the  layer
        optimisation  framework,  and  the  source  code  by  which  it  is  implemented
        were  co-developed  in equal  parts  by  \mbox{Krishnakumar Gopalakrishnan}  and
        \mbox{Ian  D.\  Campbell}.  \mbox{Krishnakumar  Gopalakrishnan}  was  the  major
        contributor  to  the  development  of  the binary  search  while  \mbox{Ian  D.\
            Campbell} was the  major contributor in the analysis  of results. \mbox{Parvathy
        Chittur Subramanianprasad}  (MSc student, Queen  Mary University of  London) was
        instrumental in  developing the analytical  expression for the  maximum possible
    number of layers~$n_\text{max}$.}%
}\label{ch:modelbaseddesign}
\ChapFrame

\vspace*{-1em}
\renewcommand{\baselinestretch}{1.0}\normalsize % for toc
\startcontents[chapters]
\printcontents[chapters]{}{1}{\setcounter{tocdepth}{1}}
\setstretch{1.348361657291667} % golden-ratio stretch (1.2 x 1.348 = 1.618)

% \bigskip
\vfill
\pagebreak

\glsunset{xeV}


\section[Introduction]{Introduction}\label{sec:layeroptintro}
\input{chapters/layer_opt/intro_layer_opt}

\section{Energy/Power Trade-off in Pouch Cells by Layer Selection}\label{sec:energypowertradeoff}
\input{chapters/layer_opt/energy_power_tradeoff}

\section{Scope and Context within \glsfmtshort{xeV} Powertrain}
\input{chapters/layer_opt/scope_drivetrain}

\section{Enhancements/Modifications to Standard \glsfmtshort{dfn} Model}\label{sec:numericalenhancements}
\input{chapters/layer_opt/numerical_enhancements}

\section{Computational Framework}\label{sec:layeroptframework}
\input{chapters/layer_opt/framework}

\section{Results and Discussion}\label{sec:resultslayeropt}
\input{chapters/layer_opt/results}

\section{Conclusions}

A methodology to design the number of layers within pouch cells so as to
maximise their energy density whilst simultaneously being capable of high
charge-acceptance rate at the plating boundary has been developed. The proposed
methodology may be employed by designers of \gls{xeV} powertrains to mitigate
the inefficiencies that plague the iterative experimental designs for each
powertrain in a vehicle manufacturer's product portfolio.  The methodology
discussed in this chapter also paves the way to incorporate common module
designs for packs through layer optimisation. By suitably changing the number of
layers within a cell, while keeping the exterior module geometry unchanged,
shall lead to cost reduction and lower the time for derivative product designs.

\addlines[0.5]
A mathematical procedure to adapt the standard galvanostatic-driven \gls{p2d}
model to accept direct power inputs has been elucidated. This reformulation is
based on a quadratic boundary condition for the discretised solid-phase charge
conservation equation, coupled with a positivity constraint on the cell's
terminal voltage and an algebraic residual equation. A key numerical aspect of
the \glspl{pbm} used in this work is the incorporation of a hybrid
pseudo-spectral/Finite Volume scheme, wherein the solid phase diffusion equation
is solved in the Chebyshev domain while a finite volume discretisation is
employed for rest of the axial-direction \glspl{pde}. This hybrid combination
helps to achieve fast simulation runs without sacrificing model fidelity.

The methodology proposed here for pouch cells may be considered for application
to cells of other form factors with suitable modifications. While considering
the cylindrical form factor, instead of a discrete set of layers, the electrode
sheets are wound into one continuous jelly roll and placed inside a cylindrical
can.  A suitable mathematical curve such as the logarithmic spiral could be
employed wherein the number of turns could be used as a proxy for the layer
count. The optimisation problem is also simplified to a certain extent since the
cylindrical can's dimensions have already been standardised across the
automotive battery industry. However, certain additional challenges could arise
in directly translating the assumptions of the cooling methodology employed. For
instance, a three dimensional discretisation of the thermal grid is not uncommon
for cylindrical form factors, which requires careful considerations for
bidirectionally coupling to the electrochemical model. Any differences in power
densities at the interior of the core versus the extremities of the electrode
spirals need to be accounted for.

This concludes the design-oriented aspects of the thesis. Next,
\cref{ch:improveddra} deals with the aspects of analysing the computational
bottlenecks present in one of the notable \glspl{rom} and proposes a mitigation
strategy for them.



% The outcome is that a ready to  use tool is made available to validate empirical
% layer choices.  In the  absence of  access to  cell manufacturing  facilities to
% confirm and test  the layer Immediate adoption  in industry, this is  the best I
% can do. Needs to be prototyped and tested.

\glsreset{xeV}

